\thetitle

### 2021-03-01 writing/rational.tex
refoctor
justify rational programmer section

Code migration is the raison d'\^etre of Typed Racket, which is why the community
uses ``migratory typing'' instead of gradual typing. Roughly speaking a
programmer uses a gradually typed language to incrementally add types to a code
base. From this perspective, a code base consists of some number of components,
say modules, and a programmer converts a component at a time from its untyped
state to a typed one.

%% MF: I don't think we should honor the nonsense of ``true gt.''  We
% are it as much as anyone!
% 
% \footnote{The method can be adapted to so-called true gradual typing
% systems~\cite{svcb-snapl-2015} in a straightforward manner.}

Based on this idea, migratory paths offer a concise setting for
evaluating the effectiveness of a rational programmer for dealing with
the interesting debugging scenarios from section ~\ref{sec:mutate}.
Grouping debugging scenarios by mutation yields a set of programs that
all contain the same mistake but differ in their type annotations.  In
the setting of migratory typing, each scenario differs from another in
which components are typed. Formally, a program $\system$ is a set of
components $\set{\component}$, and a scenario $\conf$ is the typed
subset of $\set{\component}$ . Hence, the scenarios of $\system$ are
ordered by the subset relation and, naturally, form a lattice
$\lattice{\system}$ with $2^{\size{\system}}$
elements.\footnote{\citet{tfgnvf-popl-2016} use this lattice for
performance evaluations.} The bottom scenario of $\lattice{\system}$
is always $\emptyset$ and the top one is $\system$ itself. In between
are all the mixed-typed scenarios.

In this context, the actions of the rational programmer map to an
ascending chain---dubbed a \emph{trail}---in $\lattice{\system}$.  The
starting scenario $\conf_0$ of a trail---also referred to as the
\emph{root} (of the trail)---is the initial debugging scenario; only
interesting (as described in~\ref{sec:mutate}) debugging scenarios
serve as roots.  Intuitively, starting from such a root, the rational
programmer steps along a trail, and at each step, reduces the current
scenario to an ``easier'' next scenario to debug.  The easiest
scenarios are those that the type checker rejects outright, because a
type error pinpoints an impedance mismatch between the broken code of
one component and its typed interface to others. At this point, the
programmer has identified the broken component.  If the scenario
type-checks, the programmers runs the program until it raises a run
time error.  The rational programmer uses the error's blame
assignment\footnote{Blame can result either directly from a run-time
check or, in the absence of blame, from an exception that can be
traced back to the component that causes it with the help of the stack
trace.} to decide which component(s) to equip with types. This choice
constructs scenario $\conf_{i+1}$ from $\conf_{i}$ and thus creates a trail.

% The successor of a scenario on a trail has more typed component;
% formally, configuration $\conf_{i+1}$ on the trail differs from its
% predecessor $\conf_{i}$ in that more components are typed.

The first three subsections specialize the trail-scenario idea to the
three semantics of gradual typing. Once this machinery is in place,
the fourth subsection presents a reformulation of the informal
question on page~1 as precise experimental research questions. The
final subsection adds a notion of programmer effort to the
experimental setup.

@@

\rsub{natural}   {Modes of the Natural Rational Programmer} 
\rsub{transient} {Modes of the Transient Rational Programmer} 
\rsub{erasure}   {Modes of the Erasure Rational Programmer} 
\rsub{experiment}{Experimental Process} 
\rsub{effort}    {Programmer Effort} \label{subsec:effort}

>>>>

%% -----------------------------------------------------------------------------

Section~\ref{sec:why-rational} explains, and illustrates with an example, how a
migratory type setting helps find the source of an impedance mismatch. Roughly
speaking, it encourages the rational programmer equip a module with types if it
is blamed in an error message. A sequence of such steps make up a portion of the
migratory path in the lattice of type migration~\cite{tfgnvf-popl-2016} (see
sec.~\ref{sub:stuff} for details).

Unfortunately, the rational programmer receives different kinds of information
and thus may construct different paths. As section~\ref{sec:landscape} shows,
the three different semantics deal with the same program in three different
ways, and the error messages point to different sources of a problem.

Hence the research problem is how to view the actions of a rational programmer
comparable across all semantics. Sections~\ref{sub:natural}
through~\ref{sub:erasure} describe the various rational programmers and explain
to what extent their actions are comparable.

@@

\rsub{stuff}     {Stuff}

\rsub{natural}   {The Natural Rational Programmer} 
\rsub{transient} {The Transient Rational Programmer} 
\rsub{erasure}   {The Erasure Rational Programmer}
